---
title: "Gueguen et al 2021 dataset"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(plyr)
library(dplyr)
library(patchwork)
library(cowplot)
library(ggplot2)
library(data.table)
library(tidyverse)
library(harmony)
library(org.Hs.eg.db)
library(ggrepel)
library(scater)
library(tidyverse)
library(cowplot)
library(Matrix.utils)
library(edgeR)
library(magrittr)
library(Matrix)
library(purrr)
library(reshape2)
library(S4Vectors)
library(tibble)
library(SingleCellExperiment)
library(pheatmap)
# library(apeglm)
library(png)
library(DESeq2)
library(RColorBrewer)
library(clusterProfiler)
library(sctransform)
library(gridExtra)
library(ReactomePA)
```


Downloaded dataset from: https://ncbi.xyz/geo/query/acc.cgi?acc=GSE162500

## Tumour infiltrating CD8+ lymphocytes in NSCLC: P58

```{r}
p58.counts <- Read10X("/stopgap/donglab/ling/R/megat/P58_Tumor_filtered_feature_bc_matrix") 
p58 <- CreateSeuratObject(counts = p58.counts, min.cells = 3, min.features = 200, project = "P58 TILs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p58@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p58@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p58@assays[["RNA"]])
p58 <- AddMetaData(object = p58, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p58, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p58, feature1="nCount_RNA", feature2="nFeature_RNA")

p58 <- subset(x = p58, subset = nFeature_RNA > 200 & nFeature_RNA < 4000 & percent.mito >  -Inf & percent.mito < 0.25 )

p58 <- NormalizeData(object = p58, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p58 <- FindVariableFeatures(p58, selection.method = "mean.var.plot")

# Scaling RNA data
p58 <- ScaleData(p58, features = VariableFeatures(p58))

# Run PCA
p58 <- RunPCA(p58, features = VariableFeatures(p58))

# Select top PCs for clustering
p58 <- FindNeighbors(p58, reduction = "pca", dims = 1:30, verbose=FALSE)
p58 <- FindClusters(p58, resolution = 0.5, verbose = FALSE)
p58 <- RunUMAP(p58, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p58, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Tumour infiltrating CD8+ lymphocytes in NSCLC: P60

```{r}
p60.counts <- Read10X("/stopgap/donglab/ling/R/megat/P60_Tumor_filtered_feature_bc_matrix") 
p60 <- CreateSeuratObject(counts = p60.counts, min.cells = 3, min.features = 200, project = "P60 TILs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p60@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p60@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p60@assays[["RNA"]])
p60 <- AddMetaData(object = p60, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p60, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p60, feature1="nCount_RNA", feature2="nFeature_RNA")

p60 <- subset(x = p60, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.25 )

p60 <- NormalizeData(object = p60, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p60 <- FindVariableFeatures(p60, selection.method = "mean.var.plot")

# Scaling RNA data
p60 <- ScaleData(p60, features = VariableFeatures(p60))

# Run PCA
p60 <- RunPCA(p60, features = VariableFeatures(p60))

# Select top PCs for clustering
p60 <- FindNeighbors(p60, reduction = "pca", dims = 1:30, verbose=FALSE)
p60 <- FindClusters(p60, resolution = 0.5, verbose = FALSE)
p60 <- RunUMAP(p60, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p60, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Tumour infiltrating CD8+ lymphocytes in NSCLC: P61

```{r}
p61.counts <- Read10X("/stopgap/donglab/ling/R/megat/P61_Tumor_filtered_feature_bc_matrix") 
p61 <- CreateSeuratObject(counts = p61.counts, min.cells = 3, min.features = 200, project = "P61 TILs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p61@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p61@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p61@assays[["RNA"]])
p61 <- AddMetaData(object = p61, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p61, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p61, feature1="nCount_RNA", feature2="nFeature_RNA")

p61 <- subset(x = p61, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.25 )

p61 <- NormalizeData(object = p61, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p61 <- FindVariableFeatures(p61, selection.method = "mean.var.plot")

# Scaling RNA data
p61 <- ScaleData(p61, features = VariableFeatures(p61))

# Run PCA
p61 <- RunPCA(p61, features = VariableFeatures(p61))

# Select top PCs for clustering
p61 <- FindNeighbors(p61, reduction = "pca", dims = 1:30, verbose=FALSE)
p61 <- FindClusters(p61, resolution = 0.5, verbose = FALSE)
p61 <- RunUMAP(p61, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p61, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Integrating subsets
```{r}
tils <- merge(p58, y=c(p60, p61))

tils <- NormalizeData(object = tils, normalization.method = "LogNormalize", scale.factor = 10000, verbose=FALSE)

# Select the top 1000 most variable features
tils <- FindVariableFeatures(tils, selection.method = "mean.var.plot", verbose=FALSE)

# Scaling RNA data
tils <- ScaleData(tils, features = VariableFeatures(tils), verbose=FALSE)

# Run PCA
tils <- RunPCA(tils, features = VariableFeatures(tils), verbose=FALSE)

# Select top PCs for clustering
tils <- FindNeighbors(tils, reduction = "pca", dims = 1:30, verbose=FALSE)
tils <- FindClusters(tils, resolution = 0.5, verbose = FALSE)
tils <- RunUMAP(tils, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
pca <- DimPlot(tils, reduction="pca", group.by="orig.ident")
umap_cluster <- DimPlot(tils, reduction="umap", group.by = "seurat_clusters")
umap_patient <- DimPlot(tils, reduction="umap", group.by = "orig.ident")

grid.arrange(umap_cluster, umap_patient, pca, nrow=2)

# which genes contribute to PCA loadings
VizDimLoadings(tils, dims = 1:2, reduction = "pca")
```

### Harmony integration

```{r}

tils <- tils %>%
    RunHarmony("orig.ident", plot_convergence = TRUE)

harmony_embeddings <- Embeddings(tils, 'harmony')

p1 <- DimPlot(object = tils, reduction = "harmony", group.by = "orig.ident")
p2 <- VlnPlot(object = tils, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
p1 + p2

p3 <- DimPlot(object = tils, reduction = "pca", group.by="orig.ident")
p3 + p1

tils <- tils %>%
    RunUMAP(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindNeighbors(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindClusters(resolution = 0.4, verbose=FALSE) %>%
    identity()

 
p4 <- DimPlot(tils, reduction = "umap", group.by = "orig.ident")
p5 <- DimPlot(tils, reduction = "umap", group.by = "seurat_clusters")

p4 + p5

Idents(tils) <- "seurat_clusters"
all_markers <- FindAllMarkers(tils)

saveRDS(tils, "seurat_gueguen_harmony.rds")

```

### Gene expression

```{r}
FeaturePlot(tils, features=c("XCL1", "XCL2", "ITGB3", "ITGAE"))
```
### Enriched genes

```{r}
enriched <- read.csv("/stopgap/donglab/ling/R/megat/enriched_31_proteins.csv")
tils <- readRDS("/stopgap/donglab/ling/R/megat/seurat_gueguen_harmony.rds")

tils <- AddModuleScore(object = tils,
                      features = enriched, ctrl = 5,
                      name = "enriched")

FeaturePlot(tils, features=enriched[1:12,])

FeaturePlot(tils, features="enriched1") + 
    scale_colour_gradientn(colours = c("white", "grey", "blue"))

# any cell with score 0.5 or above, annotate as enriched

metadata <- tils@meta.data

ggplot(metadata, aes(x=metadata$enriched1)) + geom_density()

metadata$enriched <- ifelse(metadata$enriched1 > 0.25, "enriched", "not enriched")
table(metadata$enriched)

tils$enriched <- metadata$enriched

DefaultAssay(tils) <- "RNA"
Idents(tils) <- "enriched"

enriched_markers <- FindMarkers(tils, ident.1="enriched", ident.2="not enriched")

# re-scale data for heatmap
tils <- ScaleData(tils, features = rownames(tils))

sig <- enriched_markers[enriched_markers$p_val_adj < 0.05,]
sig <- rownames_to_column(sig, var="gene")
sig_enriched <- sig[sig$avg_log2FC>1,]
DoHeatmap(tils, features=sig_enriched$gene) + scale_fill_gradientn(colors = c("blue", "white", "red"))
DoHeatmap(tils, features=enriched$gene) + scale_fill_gradientn(colors = c("blue", "white", "red"))
```
### Reactome pathway analysis
```{r}
genes <- sig[sig$avg_log2FC > 0,]

genes$Entrez.Gene <- mapIds(org.Hs.eg.db, keys=as.character(genes$gene), keytype="SYMBOL", column="ENTREZID")
# Keep only rows from table without NAs
genes <- genes[is.na(genes$Entrez.Gene)==FALSE,]
# Remove duplicated entries
genes <- genes[!duplicated(genes$Entrez.Gene),]

# Change Entrez IDs from numbers to characters
geneset <- as.character(genes$Entrez.Gene)

# Reactome over-representation test
x <- enrichPathway(gene=geneset, pvalueCutoff = 0.01, qvalueCutoff = 0.05, readable=TRUE)

dim(x)

reactome <- data.frame(x$ID, x$Description, x$p.adjust, x$geneID)

# split genes for each Reactome pathway into separate row
reactome <- reactome %>% mutate(x.geneID = strsplit(as.character(x.geneID), "/")) %>% unnest(x.geneID)

# enriched gene lists
enriched_105 <- read.csv("/stopgap/donglab/ling/R/megat/genes.csv")
enriched_209 <- read.csv("/stopgap/donglab/ling/R/megat/209_genes_3h.csv")
colnames(enriched_209) <- c("gene",	"CD103+_SSX-2_T_cell_clone",	
                         "CD103-_SSX-2_T_cell_clone",	"CD103+_ESO-1_T_cell_clone",
                         "CD103-_ESO-1_T_cell_clone")

# select only pathways which key genes
reactome_enriched <- reactome[reactome$x.geneID %in% enriched_105$gene |
                                  reactome$x.geneID %in% enriched_209$gene,]
reactome_enriched_pathways <- as.data.frame(table(reactome_enriched$x.Description))
colnames(reactome_enriched_pathways) <- c("Pathway","Number")

reactome_enriched_pathways <- reactome_enriched_pathways[order(-reactome_enriched_pathways$Number),]

top <- reactome_enriched_pathways[reactome_enriched_pathways$Number > 1,]


# # colour genes if on enriched list 
# reactome_enriched_genes <- genes[genes$gene %in% enriched_105$gene,]
# logfc <- reactome_enriched_genes$avg_log2FC
# names(logfc) <- reactome_enriched_genes$Entrez.Gene
# 
# cnetplot(x, foldChange = logfc, showCategory = pathways[1:5], circular = FALSE)

write.csv(reactome_enriched, "/stopgap/donglab/ling/R/megat/gueguen_upregulated_reactome_enriched_pathways.csv")
```



## Paratumour CD8+ lymphocytes in NSCLC: P60

```{r}
p60.counts <- Read10X("/stopgap/donglab/ling/R/megat/P60_Juxta_ffiltered_feature_bc_matrix") 
p60 <- CreateSeuratObject(counts = p60.counts, min.cells = 3, min.features = 200, project = "P60 Paratumour", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p60@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p60@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p60@assays[["RNA"]])
p60 <- AddMetaData(object = p60, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p60, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p60, feature1="nCount_RNA", feature2="nFeature_RNA")

p60 <- subset(x = p60, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.2 )

p60 <- NormalizeData(object = p60, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p60 <- FindVariableFeatures(p60, selection.method = "mean.var.plot")

# Scaling RNA data
p60 <- ScaleData(p60, features = VariableFeatures(p60))

# Run PCA
p60 <- RunPCA(p60, features = VariableFeatures(p60))

# Select top PCs for clustering
p60 <- FindNeighbors(p60, reduction = "pca", dims = 1:30, verbose=FALSE)
p60 <- FindClusters(p60, resolution = 0.5, verbose = FALSE)
p60 <- RunUMAP(p60, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p60, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Paratumour CD8+ lymphocytes in NSCLC: P61

```{r}
p61.counts <- Read10X("/stopgap/donglab/ling/R/megat/P61_Juxta_filtered_feature_bc_matrix") 
p61 <- CreateSeuratObject(counts = p61.counts, min.cells = 3, min.features = 200, project = "P61 Paratumour", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p61@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p61@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p61@assays[["RNA"]])
p61 <- AddMetaData(object = p61, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p61, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p61, feature1="nCount_RNA", feature2="nFeature_RNA")

p61 <- subset(x = p61, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.25 )

p61 <- NormalizeData(object = p61, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p61 <- FindVariableFeatures(p61, selection.method = "mean.var.plot")

# Scaling RNA data
p61 <- ScaleData(p61, features = VariableFeatures(p61))

# Run PCA
p61 <- RunPCA(p61, features = VariableFeatures(p61))

# Select top PCs for clustering
p61 <- FindNeighbors(p61, reduction = "pca", dims = 1:30, verbose=FALSE)
p61 <- FindClusters(p61, resolution = 0.5, verbose = FALSE)
p61 <- RunUMAP(p61, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p61, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Integrating subsets
```{r}
para <- merge(p60, y=p61)

para <- NormalizeData(object = para, normalization.method = "LogNormalize", scale.factor = 10000, verbose=FALSE)

# Select the top 1000 most variable features
para <- FindVariableFeatures(para, selection.method = "mean.var.plot", verbose=FALSE)

# Scaling RNA data
para <- ScaleData(para, features = VariableFeatures(para), verbose=FALSE)

# Run PCA
para <- RunPCA(para, features = VariableFeatures(para), verbose=FALSE)

# Select top PCs for clustering
para <- FindNeighbors(para, reduction = "pca", dims = 1:30, verbose=FALSE)
para <- FindClusters(para, resolution = 0.5, verbose = FALSE)
para <- RunUMAP(para, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
pca <- DimPlot(para, reduction="pca", group.by="orig.ident")
umap_cluster <- DimPlot(para, reduction="umap", group.by = "seurat_clusters")
umap_patient <- DimPlot(para, reduction="umap", group.by = "orig.ident")

grid.arrange(umap_cluster, umap_patient, pca, nrow=2)

# which genes contribute to PCA loadings
VizDimLoadings(para, dims = 1:2, reduction = "pca")
```

### Harmony integration

```{r}
para <- para %>%
    RunHarmony("orig.ident", plot_convergence = TRUE)

harmony_embeddings <- Embeddings(para, 'harmony')

p1 <- DimPlot(object = para, reduction = "harmony", group.by = "orig.ident")
p2 <- VlnPlot(object = para, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
p1 + p2

p3 <- DimPlot(object = para, reduction = "pca", group.by="orig.ident")
p3 + p1

para <- para %>%
    RunUMAP(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindNeighbors(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindClusters(resolution = 0.4, verbose=FALSE) %>%
    identity()

 
p4 <- DimPlot(para, reduction = "umap", group.by = "orig.ident")
p5 <- DimPlot(para, reduction = "umap", group.by = "seurat_clusters")

p4 + p5

```

### Gene expression

```{r}
enriched <- read.csv("/stopgap/donglab/ling/R/megat/genes.csv")
FeaturePlot(para, features=c("XCL1", "XCL2", "ITGB3", "ITGAE"))
```




## PMBCs CD8+ lymphocytes in NSCLC: P58

```{r}
p58.counts <- Read10X("/stopgap/donglab/ling/R/megat/P58_Blood_filtered_feature_bc_matrix") 
p58 <- CreateSeuratObject(counts = p58.counts, min.cells = 3, min.features = 200, project = "P58 PBMCs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p58@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p58@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p58@assays[["RNA"]])
p58 <- AddMetaData(object = p58, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p58, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p58, feature1="nCount_RNA", feature2="nFeature_RNA")

p58 <- subset(x = p58, subset = nFeature_RNA > 200 & nFeature_RNA < 4000 & percent.mito >  -Inf & percent.mito < 0.2 )

p58 <- NormalizeData(object = p58, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p58 <- FindVariableFeatures(p58, selection.method = "mean.var.plot")

# Scaling RNA data
p58 <- ScaleData(p58, features = VariableFeatures(p58))

# Run PCA
p58 <- RunPCA(p58, features = VariableFeatures(p58))

# Select top PCs for clustering
p58 <- FindNeighbors(p58, reduction = "pca", dims = 1:30, verbose=FALSE)
p58 <- FindClusters(p58, resolution = 0.5, verbose = FALSE)
p58 <- RunUMAP(p58, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p58, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## PBMC CD8+ lymphocytes in NSCLC: P60

```{r}
p60.counts <- Read10X("/stopgap/donglab/ling/R/megat/P60_Blood_filtered_feature_bc_matrix") 
p60 <- CreateSeuratObject(counts = p60.counts, min.cells = 3, min.features = 200, project = "P60 PBMCs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p60@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p60@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p60@assays[["RNA"]])
p60 <- AddMetaData(object = p60, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p60, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p60, feature1="nCount_RNA", feature2="nFeature_RNA")

p60 <- subset(x = p60, subset = nFeature_RNA > 200 & nFeature_RNA < 4500 & percent.mito >  -Inf & percent.mito < 0.15 )

p60 <- NormalizeData(object = p60, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p60 <- FindVariableFeatures(p60, selection.method = "mean.var.plot")

# Scaling RNA data
p60 <- ScaleData(p60, features = VariableFeatures(p60))

# Run PCA
p60 <- RunPCA(p60, features = VariableFeatures(p60))

# Select top PCs for clustering
p60 <- FindNeighbors(p60, reduction = "pca", dims = 1:30, verbose=FALSE)
p60 <- FindClusters(p60, resolution = 0.5, verbose = FALSE)
p60 <- RunUMAP(p60, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p60, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## PBMCs CD8+ lymphocytes in NSCLC: P61

```{r}
p61.counts <- Read10X("/stopgap/donglab/ling/R/megat/P61_Blood_filtered_feature_bc_matrix") 
p61 <- CreateSeuratObject(counts = p61.counts, min.cells = 3, min.features = 200, project = "P61 PBMCs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p61@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p61@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p61@assays[["RNA"]])
p61 <- AddMetaData(object = p61, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p61, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p61, feature1="nCount_RNA", feature2="nFeature_RNA")

p61 <- subset(x = p61, subset = nFeature_RNA > 200 & nFeature_RNA < 4500 & percent.mito >  -Inf & percent.mito < 0.25 )

p61 <- NormalizeData(object = p61, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p61 <- FindVariableFeatures(p61, selection.method = "mean.var.plot")

# Scaling RNA data
p61 <- ScaleData(p61, features = VariableFeatures(p61))

# Run PCA
p61 <- RunPCA(p61, features = VariableFeatures(p61))

# Select top PCs for clustering
p61 <- FindNeighbors(p61, reduction = "pca", dims = 1:30, verbose=FALSE)
p61 <- FindClusters(p61, resolution = 0.5, verbose = FALSE)
p61 <- RunUMAP(p61, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p61, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Integrating subsets
```{r}
pbmc <- merge(p58, y=c(p60, p61))

pbmc <- NormalizeData(object = pbmc, normalization.method = "LogNormalize", scale.factor = 10000, verbose=FALSE)

# Select the top 1000 most variable features
pbmc <- FindVariableFeatures(pbmc, selection.method = "mean.var.plot", verbose=FALSE)

# Scaling RNA data
pbmc <- ScaleData(pbmc, features = VariableFeatures(pbmc), verbose=FALSE)

# Run PCA
pbmc <- RunPCA(pbmc, features = VariableFeatures(pbmc), verbose=FALSE)

# Select top PCs for clustering
pbmc <- FindNeighbors(pbmc, reduction = "pca", dims = 1:30, verbose=FALSE)
pbmc <- FindClusters(pbmc, resolution = 0.5, verbose = FALSE)
pbmc <- RunUMAP(pbmc, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
pca <- DimPlot(pbmc, reduction="pca", group.by="orig.ident")
umap_cluster <- DimPlot(pbmc, reduction="umap", group.by = "seurat_clusters")
umap_patient <- DimPlot(pbmc, reduction="umap", group.by = "orig.ident")

grid.arrange(umap_cluster, umap_patient, pca, nrow=2)

# which genes contribute to PCA loadings
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```

### Harmony integration

```{r}
pbmc <- pbmc %>%
    RunHarmony("orig.ident", plot_convergence = TRUE)

harmony_embeddings <- Embeddings(pbmc, 'harmony')

p1 <- DimPlot(object = pbmc, reduction = "harmony", group.by = "orig.ident")
p2 <- VlnPlot(object = pbmc, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
p1 + p2

p3 <- DimPlot(object = pbmc, reduction = "pca", group.by="orig.ident")
p3 + p1

pbmc <- pbmc %>%
    RunUMAP(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindNeighbors(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindClusters(resolution = 0.4, verbose=FALSE) %>%
    identity()

 
p4 <- DimPlot(pbmc, reduction = "umap", group.by = "orig.ident")
p5 <- DimPlot(pbmc, reduction = "umap", group.by = "seurat_clusters")

p4 + p5

```

### Gene expression

```{r}
enriched <- read.csv("/stopgap/donglab/ling/R/megat/genes.csv")
FeaturePlot(pbmc, features=c("XCL1", "XCL2", "ITGB3", "ITGAE"))
```



