---
title: "Gueguen et al 2021 dataset"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(plyr)
library(dplyr)
library(patchwork)
library(cowplot)
library(ggplot2)
library(data.table)
library(tidyverse)
library(harmony)
library(org.Hs.eg.db)
library(ggrepel)
library(scater)
library(tidyverse)
library(cowplot)
library(Matrix.utils)
library(edgeR)
library(magrittr)
library(Matrix)
library(purrr)
library(reshape2)
library(S4Vectors)
library(tibble)
library(SingleCellExperiment)
library(pheatmap)
# library(apeglm)
library(png)
library(DESeq2)
library(RColorBrewer)
library(clusterProfiler)
library(sctransform)
library(gridExtra)
```


Downloaded dataset from: https://ncbi.xyz/geo/query/acc.cgi?acc=GSE162500

## Tumour infiltrating CD8+ lymphocytes in NSCLC: P58

```{r}
p58.counts <- Read10X("/stopgap/donglab/ling/R/megat/P58_Tumor_filtered_feature_bc_matrix") 
p58 <- CreateSeuratObject(counts = p58.counts, min.cells = 3, min.features = 200, project = "P58 TILs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p58@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p58@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p58@assays[["RNA"]])
p58 <- AddMetaData(object = p58, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p58, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p58, feature1="nCount_RNA", feature2="nFeature_RNA")

p58 <- subset(x = p58, subset = nFeature_RNA > 200 & nFeature_RNA < 4000 & percent.mito >  -Inf & percent.mito < 0.25 )

p58 <- NormalizeData(object = p58, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p58 <- FindVariableFeatures(p58, selection.method = "mean.var.plot")

# Scaling RNA data
p58 <- ScaleData(p58, features = VariableFeatures(p58))

# Run PCA
p58 <- RunPCA(p58, features = VariableFeatures(p58))

# Select top PCs for clustering
p58 <- FindNeighbors(p58, reduction = "pca", dims = 1:30, verbose=FALSE)
p58 <- FindClusters(p58, resolution = 0.5, verbose = FALSE)
p58 <- RunUMAP(p58, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p58, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Tumour infiltrating CD8+ lymphocytes in NSCLC: P60

```{r}
p60.counts <- Read10X("/stopgap/donglab/ling/R/megat/P60_Tumor_filtered_feature_bc_matrix") 
p60 <- CreateSeuratObject(counts = p60.counts, min.cells = 3, min.features = 200, project = "P60 TILs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p60@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p60@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p60@assays[["RNA"]])
p60 <- AddMetaData(object = p60, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p60, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p60, feature1="nCount_RNA", feature2="nFeature_RNA")

p60 <- subset(x = p60, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.25 )

p60 <- NormalizeData(object = p60, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p60 <- FindVariableFeatures(p60, selection.method = "mean.var.plot")

# Scaling RNA data
p60 <- ScaleData(p60, features = VariableFeatures(p60))

# Run PCA
p60 <- RunPCA(p60, features = VariableFeatures(p60))

# Select top PCs for clustering
p60 <- FindNeighbors(p60, reduction = "pca", dims = 1:30, verbose=FALSE)
p60 <- FindClusters(p60, resolution = 0.5, verbose = FALSE)
p60 <- RunUMAP(p60, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p60, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Tumour infiltrating CD8+ lymphocytes in NSCLC: P61

```{r}
p61.counts <- Read10X("/stopgap/donglab/ling/R/megat/P61_Tumor_filtered_feature_bc_matrix") 
p61 <- CreateSeuratObject(counts = p61.counts, min.cells = 3, min.features = 200, project = "P61 TILs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p61@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p61@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p61@assays[["RNA"]])
p61 <- AddMetaData(object = p61, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p61, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p61, feature1="nCount_RNA", feature2="nFeature_RNA")

p61 <- subset(x = p61, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.25 )

p61 <- NormalizeData(object = p61, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p61 <- FindVariableFeatures(p61, selection.method = "mean.var.plot")

# Scaling RNA data
p61 <- ScaleData(p61, features = VariableFeatures(p61))

# Run PCA
p61 <- RunPCA(p61, features = VariableFeatures(p61))

# Select top PCs for clustering
p61 <- FindNeighbors(p61, reduction = "pca", dims = 1:30, verbose=FALSE)
p61 <- FindClusters(p61, resolution = 0.5, verbose = FALSE)
p61 <- RunUMAP(p61, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p61, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Integrating subsets
```{r}
tils <- merge(p58, y=c(p60, p61))

tils <- NormalizeData(object = tils, normalization.method = "LogNormalize", scale.factor = 10000, verbose=FALSE)

# Select the top 1000 most variable features
tils <- FindVariableFeatures(tils, selection.method = "mean.var.plot", verbose=FALSE)

# Scaling RNA data
tils <- ScaleData(tils, features = VariableFeatures(tils), verbose=FALSE)

# Run PCA
tils <- RunPCA(tils, features = VariableFeatures(tils), verbose=FALSE)

# Select top PCs for clustering
tils <- FindNeighbors(tils, reduction = "pca", dims = 1:30, verbose=FALSE)
tils <- FindClusters(tils, resolution = 0.5, verbose = FALSE)
tils <- RunUMAP(tils, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
pca <- DimPlot(tils, reduction="pca", group.by="orig.ident")
umap_cluster <- DimPlot(tils, reduction="umap", group.by = "seurat_clusters")
umap_patient <- DimPlot(tils, reduction="umap", group.by = "orig.ident")

grid.arrange(umap_cluster, umap_patient, pca, nrow=2)

# which genes contribute to PCA loadings
VizDimLoadings(tils, dims = 1:2, reduction = "pca")
```

### Harmony integration

```{r}

tils <- tils %>%
    RunHarmony("orig.ident", plot_convergence = TRUE)

harmony_embeddings <- Embeddings(tils, 'harmony')

p1 <- DimPlot(object = tils, reduction = "harmony", group.by = "orig.ident")
p2 <- VlnPlot(object = tils, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
p1 + p2

p3 <- DimPlot(object = tils, reduction = "pca", group.by="orig.ident")
p3 + p1

tils <- tils %>%
    RunUMAP(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindNeighbors(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindClusters(resolution = 0.4, verbose=FALSE) %>%
    identity()

 
p4 <- DimPlot(tils, reduction = "umap", group.by = "orig.ident")
p5 <- DimPlot(tils, reduction = "umap", group.by = "seurat_clusters")

p4 + p5

```

### Gene expression

```{r}
FeaturePlot(tils, features=c("XCL1", "XCL2", "ITGB3", "ITGAE"))
```
### Enriched genes

```{r}
enriched <- read.csv("/stopgap/donglab/ling/R/megat/genes.csv")

tils <- AddModuleScore(object = tils,
                      features = enriched, ctrl = 5,
                      name = "enriched")

FeaturePlot(tils, features=enriched[1:12,])

FeaturePlot(tils, features="enriched1")
```



## Paratumour CD8+ lymphocytes in NSCLC: P60

```{r}
p60.counts <- Read10X("/stopgap/donglab/ling/R/megat/P60_Juxta_ffiltered_feature_bc_matrix") 
p60 <- CreateSeuratObject(counts = p60.counts, min.cells = 3, min.features = 200, project = "P60 Paratumour", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p60@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p60@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p60@assays[["RNA"]])
p60 <- AddMetaData(object = p60, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p60, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p60, feature1="nCount_RNA", feature2="nFeature_RNA")

p60 <- subset(x = p60, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.2 )

p60 <- NormalizeData(object = p60, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p60 <- FindVariableFeatures(p60, selection.method = "mean.var.plot")

# Scaling RNA data
p60 <- ScaleData(p60, features = VariableFeatures(p60))

# Run PCA
p60 <- RunPCA(p60, features = VariableFeatures(p60))

# Select top PCs for clustering
p60 <- FindNeighbors(p60, reduction = "pca", dims = 1:30, verbose=FALSE)
p60 <- FindClusters(p60, resolution = 0.5, verbose = FALSE)
p60 <- RunUMAP(p60, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p60, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Paratumour CD8+ lymphocytes in NSCLC: P61

```{r}
p61.counts <- Read10X("/stopgap/donglab/ling/R/megat/P61_Juxta_filtered_feature_bc_matrix") 
p61 <- CreateSeuratObject(counts = p61.counts, min.cells = 3, min.features = 200, project = "P61 Paratumour", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p61@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p61@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p61@assays[["RNA"]])
p61 <- AddMetaData(object = p61, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p61, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p61, feature1="nCount_RNA", feature2="nFeature_RNA")

p61 <- subset(x = p61, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.25 )

p61 <- NormalizeData(object = p61, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p61 <- FindVariableFeatures(p61, selection.method = "mean.var.plot")

# Scaling RNA data
p61 <- ScaleData(p61, features = VariableFeatures(p61))

# Run PCA
p61 <- RunPCA(p61, features = VariableFeatures(p61))

# Select top PCs for clustering
p61 <- FindNeighbors(p61, reduction = "pca", dims = 1:30, verbose=FALSE)
p61 <- FindClusters(p61, resolution = 0.5, verbose = FALSE)
p61 <- RunUMAP(p61, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p61, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Integrating subsets
```{r}
para <- merge(p60, y=p61)

para <- NormalizeData(object = para, normalization.method = "LogNormalize", scale.factor = 10000, verbose=FALSE)

# Select the top 1000 most variable features
para <- FindVariableFeatures(para, selection.method = "mean.var.plot", verbose=FALSE)

# Scaling RNA data
para <- ScaleData(para, features = VariableFeatures(para), verbose=FALSE)

# Run PCA
para <- RunPCA(para, features = VariableFeatures(para), verbose=FALSE)

# Select top PCs for clustering
para <- FindNeighbors(para, reduction = "pca", dims = 1:30, verbose=FALSE)
para <- FindClusters(para, resolution = 0.5, verbose = FALSE)
para <- RunUMAP(para, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
pca <- DimPlot(para, reduction="pca", group.by="orig.ident")
umap_cluster <- DimPlot(para, reduction="umap", group.by = "seurat_clusters")
umap_patient <- DimPlot(para, reduction="umap", group.by = "orig.ident")

grid.arrange(umap_cluster, umap_patient, pca, nrow=2)

# which genes contribute to PCA loadings
VizDimLoadings(para, dims = 1:2, reduction = "pca")
```

### Harmony integration

```{r}
para <- para %>%
    RunHarmony("orig.ident", plot_convergence = TRUE)

harmony_embeddings <- Embeddings(para, 'harmony')

p1 <- DimPlot(object = para, reduction = "harmony", group.by = "orig.ident")
p2 <- VlnPlot(object = para, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
p1 + p2

p3 <- DimPlot(object = para, reduction = "pca", group.by="orig.ident")
p3 + p1

para <- para %>%
    RunUMAP(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindNeighbors(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindClusters(resolution = 0.4, verbose=FALSE) %>%
    identity()

 
p4 <- DimPlot(para, reduction = "umap", group.by = "orig.ident")
p5 <- DimPlot(para, reduction = "umap", group.by = "seurat_clusters")

p4 + p5

```

### Gene expression

```{r}
enriched <- read.csv("/stopgap/donglab/ling/R/megat/genes.csv")
FeaturePlot(para, features=c("XCL1", "XCL2", "ITGB3", "ITGAE"))
```




## PMBCs CD8+ lymphocytes in NSCLC: P58

```{r}
p58.counts <- Read10X("/stopgap/donglab/ling/R/megat/P58_Blood_filtered_feature_bc_matrix") 
p58 <- CreateSeuratObject(counts = p58.counts, min.cells = 3, min.features = 200, project = "P58 PBMCs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p58@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p58@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p58@assays[["RNA"]])
p58 <- AddMetaData(object = p58, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p58, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p58, feature1="nCount_RNA", feature2="nFeature_RNA")

p58 <- subset(x = p58, subset = nFeature_RNA > 200 & nFeature_RNA < 4000 & percent.mito >  -Inf & percent.mito < 0.2 )

p58 <- NormalizeData(object = p58, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p58 <- FindVariableFeatures(p58, selection.method = "mean.var.plot")

# Scaling RNA data
p58 <- ScaleData(p58, features = VariableFeatures(p58))

# Run PCA
p58 <- RunPCA(p58, features = VariableFeatures(p58))

# Select top PCs for clustering
p58 <- FindNeighbors(p58, reduction = "pca", dims = 1:30, verbose=FALSE)
p58 <- FindClusters(p58, resolution = 0.5, verbose = FALSE)
p58 <- RunUMAP(p58, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p58, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## PBMC CD8+ lymphocytes in NSCLC: P60

```{r}
p60.counts <- Read10X("/stopgap/donglab/ling/R/megat/P60_Blood_filtered_feature_bc_matrix") 
p60 <- CreateSeuratObject(counts = p60.counts, min.cells = 3, min.features = 200, project = "P60 PBMCs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p60@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p60@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p60@assays[["RNA"]])
p60 <- AddMetaData(object = p60, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p60, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p60, feature1="nCount_RNA", feature2="nFeature_RNA")

p60 <- subset(x = p60, subset = nFeature_RNA > 200 & nFeature_RNA < 4500 & percent.mito >  -Inf & percent.mito < 0.15 )

p60 <- NormalizeData(object = p60, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p60 <- FindVariableFeatures(p60, selection.method = "mean.var.plot")

# Scaling RNA data
p60 <- ScaleData(p60, features = VariableFeatures(p60))

# Run PCA
p60 <- RunPCA(p60, features = VariableFeatures(p60))

# Select top PCs for clustering
p60 <- FindNeighbors(p60, reduction = "pca", dims = 1:30, verbose=FALSE)
p60 <- FindClusters(p60, resolution = 0.5, verbose = FALSE)
p60 <- RunUMAP(p60, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p60, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## PBMCs CD8+ lymphocytes in NSCLC: P61

```{r}
p61.counts <- Read10X("/stopgap/donglab/ling/R/megat/P61_Blood_filtered_feature_bc_matrix") 
p61 <- CreateSeuratObject(counts = p61.counts, min.cells = 3, min.features = 200, project = "P61 PBMCs", assay = "RNA")

mito.genes <- grep(pattern = "^MT-", x = rownames(p61@assays[["RNA"]]), value = TRUE)
percent.mito <- Matrix::colSums(p61@assays[["RNA"]][mito.genes, ])/Matrix::colSums(p61@assays[["RNA"]])
p61 <- AddMetaData(object = p61, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = p61, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

FeatureScatter(object=p61, feature1="nCount_RNA", feature2="nFeature_RNA")

p61 <- subset(x = p61, subset = nFeature_RNA > 200 & nFeature_RNA < 4500 & percent.mito >  -Inf & percent.mito < 0.25 )

p61 <- NormalizeData(object = p61, normalization.method = "LogNormalize", scale.factor = 10000)

# Select the top 1000 most variable features
p61 <- FindVariableFeatures(p61, selection.method = "mean.var.plot")

# Scaling RNA data
p61 <- ScaleData(p61, features = VariableFeatures(p61))

# Run PCA
p61 <- RunPCA(p61, features = VariableFeatures(p61))

# Select top PCs for clustering
p61 <- FindNeighbors(p61, reduction = "pca", dims = 1:30, verbose=FALSE)
p61 <- FindClusters(p61, resolution = 0.5, verbose = FALSE)
p61 <- RunUMAP(p61, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
umap_cluster <- DimPlot(p61, reduction="umap", group.by = "seurat_clusters")

umap_cluster
```

## Integrating subsets
```{r}
pbmc <- merge(p58, y=c(p60, p61))

pbmc <- NormalizeData(object = pbmc, normalization.method = "LogNormalize", scale.factor = 10000, verbose=FALSE)

# Select the top 1000 most variable features
pbmc <- FindVariableFeatures(pbmc, selection.method = "mean.var.plot", verbose=FALSE)

# Scaling RNA data
pbmc <- ScaleData(pbmc, features = VariableFeatures(pbmc), verbose=FALSE)

# Run PCA
pbmc <- RunPCA(pbmc, features = VariableFeatures(pbmc), verbose=FALSE)

# Select top PCs for clustering
pbmc <- FindNeighbors(pbmc, reduction = "pca", dims = 1:30, verbose=FALSE)
pbmc <- FindClusters(pbmc, resolution = 0.5, verbose = FALSE)
pbmc <- RunUMAP(pbmc, dims = 1:30, verbose=FALSE)

# Projecting singlet identities on UMAP visualization
pca <- DimPlot(pbmc, reduction="pca", group.by="orig.ident")
umap_cluster <- DimPlot(pbmc, reduction="umap", group.by = "seurat_clusters")
umap_patient <- DimPlot(pbmc, reduction="umap", group.by = "orig.ident")

grid.arrange(umap_cluster, umap_patient, pca, nrow=2)

# which genes contribute to PCA loadings
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```

### Harmony integration

```{r}
pbmc <- pbmc %>%
    RunHarmony("orig.ident", plot_convergence = TRUE)

harmony_embeddings <- Embeddings(pbmc, 'harmony')

p1 <- DimPlot(object = pbmc, reduction = "harmony", group.by = "orig.ident")
p2 <- VlnPlot(object = pbmc, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
p1 + p2

p3 <- DimPlot(object = pbmc, reduction = "pca", group.by="orig.ident")
p3 + p1

pbmc <- pbmc %>%
    RunUMAP(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindNeighbors(reduction = "harmony", dims = 1:20, verbose=FALSE) %>%
    FindClusters(resolution = 0.4, verbose=FALSE) %>%
    identity()

 
p4 <- DimPlot(pbmc, reduction = "umap", group.by = "orig.ident")
p5 <- DimPlot(pbmc, reduction = "umap", group.by = "seurat_clusters")

p4 + p5

```

### Gene expression

```{r}
enriched <- read.csv("/stopgap/donglab/ling/R/megat/genes.csv")
FeaturePlot(pbmc, features=c("XCL1", "XCL2", "ITGB3", "ITGAE"))
```



